---
title: "LTER material legacy synthesis"
author: "Kai L. Kopecky"
date: "2024-08-20"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(glmmTMB)
library(ggeffects)
library(broom.mixed)
```

# Exploratory looks at data from LTER sites

## Konza Prairie
```{r}
# Explore relationship of previous year dead biomass on current year productivity
# Should not group forbs and grass; grass could be considered the foundation species, and forbs neither produce legacies in the same way, nor do are they annuals

# Look at relationship between dead biomass and biomass of forbs and woody 

# potential relationships between dead biomass and abiotic conditions (soil temperature, soil moisture, N content)

# Look into % cover data as well

# Watershed = fire treatment: 004 means burned every four years, 020 = every 20 years, b = statistical replicate
# Soiltype: tu = lowland, thick, moist; fl = upland, thin, rocky, drier

# Consider trying a running mean approach

knz <- read_csv("Data/Konza Prairie/PAB011.csv") %>% 
  clean_names() %>% 
  filter(pryrdead >=0,
         lvgrass != "NA")

hist(knz$lvgrass)
hist(log(knz$lvgrass))

hist(knz$pryrdead)
hist(log(knz$pryrdead))

# Exploratory graph
ggplot(knz, aes(x = pryrdead, y = lvgrass, color = watershed)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", 
              aes(color = watershed)) +
  #scale_x_continuous(expand = c(0.01,1)) +
 # scale_y_continuous(expand = c(0.01,1)) +
  labs(x = "Lagged dead biomass",
       y = "Current year biomass") +
  theme_classic()

# Forbs ~ previous year
ggplot(knz, aes(x = pryrdead, y = forbs, color = watershed)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", 
              aes(color = watershed)) +
  labs(x = "Lagged dead biomass",
       y = "Current year forb biomass") +
  theme_classic()

ggplot(knz, aes(x = pryrdead, y = woody, color = watershed)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", 
              aes(color = watershed)) +
  labs(x = "Lagged dead biomass",
       y = "Current year forb biomass") +
  theme_classic()

```

Mixed effects models
```{r}
# glmm of total live biomass ~ previous year dead biomass
knz.glmm <- glmmTMB(lvgrass ~ pryrdead + (1|transect),
                    data = knz)
summary(knz.glmm)

predictions.knz <- ggpredict(knz.glmm, terms = ~ pryrdead + (1|transect))
plot(predictions.knz)
predictions.knz <- as.data.frame(predictions.knz)

# Model visualization
ggplot()+
  geom_point(data = knz, 
             aes(x = pryrdead, y = lvgrass),
             alpha = 0.6,
             size = 2) +
  geom_ribbon(data = predictions.knz, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              fill = "#3B4F8E",
              alpha =0.5)+
  geom_line(data = predictions.knz, 
            aes(x = x , y = predicted, group = group)) +
  scale_x_continuous(expand = c(0.01,1)) +
  scale_y_continuous(#limits = c(0,160), 
                     expand = c(0.01,1)) +
  labs(x = "Previous year dead biomass (g/m^2)",
       y = "Current year live biomass (g/m^2)") +
  theme_classic()

# glmm of total live biomass ~ previous year dead biomass*watershed
knz.glmm <- glmmTMB(lvgrass ~ pryrdead*watershed + (1|transect),
                    data = knz)
summary(knz.glmm)

predictions.knz <- ggpredict(knz.glmm, terms = ~ pryrdead*watershed + (1|transect))
plot(predictions.knz)
predictions.knz <- as.data.frame(predictions.knz)

# Model visualization
ggplot()+
  geom_point(data = knz, 
             aes(x = pryrdead, y = lvgrass, color = watershed),
             alpha = 0.6,
             size = 2) +
  geom_ribbon(data = predictions.knz, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high, fill = group), 
              #fill = "#3B4F8E",
              alpha =0.5)+
  geom_line(data = predictions.knz, 
            aes(x = x , y = predicted, group = group, color = group)) +
  scale_x_continuous(expand = c(0.01,1)) +
  scale_y_continuous(#limits = c(0,160), 
                     expand = c(0.01,1)) +
  labs(x = "Previous year dead biomass (g/m^2)",
       y = "Current year live biomass (g/m^2)") +
  theme_classic()

```

```{r}
# Normalized biomasses
# Scaled to maximum values
knz <- knz %>% 
  mutate(pryrdead.norm = pryrdead/max(knz$pryrdead),
         live_grass.norm = lvgrass/max(knz$lvgrass))

# glmm of normalized total live biomass ~ normalized previous year dead biomass
knz.norm.glmm <- glmmTMB(live_grass.norm ~ pryrdead.norm + (1|transect),
                    data = knz)
summary(knz.glmm)

predictions.knz.norm <- ggpredict(knz.norm.glmm, terms = ~ pryrdead.norm + (1|transect))
plot(predictions.knz.norm)
predictions.knz.norm <- as.data.frame(predictions.knz.norm)

# Model visualization of normalized biomasses
ggplot()+
  geom_point(data = knz, 
             aes(x = pryrdead.norm, y = live_grass.norm),
             alpha = 0.6,
             size = 2) +
  geom_ribbon(data = predictions.knz.norm, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = predictions.knz.norm, 
            aes(x = x , y = predicted))+
  scale_x_continuous(expand = c(0,0.02),
                     limits = c(0,1.01)) +
  scale_y_continuous(expand = c(0,0.02),
                     limits = c(0,1.01)) +
  labs(x = "Normalized previous year dead biomass",
       y = "Nomralized live biomass") +
  theme_classic()
```

```{r Analyses by plot}

# Sum all biomass across plots within a year
knz.year_totals <- knz %>% 
  group_by(recyear, watershed, transect, plotnum) %>% 
  summarize(live_biomass.total = sum(total_live_biomass),
            #live_biomass.log = log(live_biomass.total + 1),
            dead_biomass.prev = sum(pryrdead),
            #dead_biomass.prev.log = log(dead_biomass.prev + 1),
            dead_biomass.total = sum(total_dead_biomass))
            #dead_biomass.log = log(dead_biomass.total + 1)) 

ggplot(knz.year_totals, aes(x = dead_biomass.total, y = live_biomass.total, color = watershed)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  scale_x_continuous(expand = c(0.01,1)) +
  scale_y_continuous(expand = c(0.01,1)) +
  labs(x = "Previous year dead biomass",
       y = "Current year live biomass") +
  theme_classic()


## glmm with total dead_biomass (previous + current year)
knz.glmm <- glmmTMB(live_biomass.total ~ dead_biomass.total + (1|transect),
                    data = knz.year_totals)
summary(knz.glmm)

predictions.knz <- ggpredict(knz.glmm, terms = ~dead_biomass.total + (1|transect))
plot(predictions.knz)
predictions.knz <- as.data.frame(predictions.knz)

# Model visualization
ggplot()+
  geom_ribbon(data = predictions.knz, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = predictions.knz, 
            aes(x = x , y = predicted))+
  geom_point(data = knz.year_totals, 
             aes(x = dead_biomass.total, y = live_biomass.total),
             alpha = 0.6,
             size = 2) +
  labs(x = "Previous + current year dead biomass",
       y = "Current year live biomass") +
  theme_classic()

## glmm with just prvious year dead biomass
knz.glmm.prev <- glmmTMB(live_biomass.total ~ dead_biomass.prev + (1|transect),
                    data = knz.year_totals)
summary(knz.glmm.prev)

predictions.knz.prev <- ggpredict(knz.glmm.prev, terms = ~dead_biomass.prev + (1|transect))
plot(predictions.knz.prev)
predictions.knz.prev <- as.data.frame(predictions.knz.prev)

# Model visualization
ggplot()+
  geom_ribbon(data = predictions.knz.prev, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = predictions.knz.prev, 
            aes(x = x , y = predicted))+
  geom_point(data = knz.year_totals, 
             aes(x = dead_biomass.prev, y = live_biomass.total),
             alpha = 0.6,
             size = 2) +
  labs(x = "Previous year dead biomass",
       y = "Current year live biomass") +
  theme_classic()


```


## Virginia Coastal Reserve
```{r Data cleaning and wrangling}

# Use only reference sites; or, also use restored sites that are >= 6 years since construction

# Plot juv denisty against day of year 

# lump seasons; conservative approach could be to limit analysis to summer-fall

# plot live oyster against dead

vcr.raw <- read_csv("Data/Virginia Coastal/Oyster_Count_Data_2022_03_10.csv")

# Filter original data for just live and dead oysters
vcr <- read_csv("Data/Virginia Coastal/Oyster_Count_Data_2022_03_10.csv") %>% 
  clean_names() %>% 
  select(site, sample, date, species, species_count) %>% 
  filter(species == c("Adult Oyster", "Box Adult Oyster", "Box Spat Oyster", "Spat Oyster")) %>% 
  mutate(year = substr(date, nchar(date) - 1, nchar(date)),
         year = paste0("20", year),
         date = mdy(date), # Convert the date column to Date format
         month = month(date), # Extract the month
         season = case_when( # Assign seasons based on month
          month %in% c(12, 1, 2) ~ "Winter",
          month %in% c(3, 4, 5) ~ "Spring",
          month %in% c(6, 7, 8) ~ "Summer",
          month %in% c(9, 10, 11) ~ "Fall",
          TRUE ~ NA_character_))


# Create summary dataframe for juvenile oyster densities 
vcr.juv_density <- vcr %>% 
  mutate(species = case_when(species == "Adult Oyster" ~ "Adult",
                             species == "Spat Oyster" ~ "Spat",
                             TRUE ~ "Dead")) %>% 
  filter(species != "Adult") %>% 
  group_by(site, species, year, season) %>% 
  summarize(mean_density = mean(species_count)) %>% 
  pivot_wider(names_from = species, values_from = mean_density) %>% 
  drop_na()

# Explore seasonal trends in juvenile density
vcr.juv_density.season <- vcr.juv_density %>% 
  group_by(season) %>% 
  summarize(agg.juv_density = sum(Spat),
            mean.juv_density = mean(Spat),
            se.juv_density = sd(Spat)/sqrt(n()))

ggplot(vcr.juv_density.season, aes(x = season, y = agg.juv_density, fill = site)) +
  geom_col(position = "dodge")

ggplot(vcr.juv_density.season, aes(x = season, y = mean.juv_density)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin = mean.juv_density - se.juv_density,
                    ymax = mean.juv_density + se.juv_density),
                    width = 0)

# Create dataframe with dead oyster density lagged one year behind juvenile density
vcr_lagged <- vcr.juv_density %>% 
  group_by(site, season) %>% 
  mutate(lagged_Dead = lag(Dead)) %>% 
  filter(lagged_Dead < 100)

## Visualizations
# Juvenile oyster density ~ dead oyster density in previous year (lagged)
ggplot(vcr_lagged, aes(x = lagged_Dead, y = Spat, color = season)) +
  geom_point(alpha = 0.5,
             aes(color = season)) +
  geom_smooth(method = "lm",
              aes(group = season)) +
  #scale_x_continuous(expand = c(0.01,0)) +
  #scale_y_continuous(expand = c(0.01,0)) +
  labs(x = "Dead oyster density",
       y = "Juvenile oyster density") +
  theme_classic()

# Normalized juvenile oyster density ~ normalized dead oyster density in previous year (lagged)
ggplot(vcr_lagged, aes(x = lagged_dead.norm, y = spat.norm)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  scale_x_continuous(expand = c(0.01,0)) +
  scale_y_continuous(expand = c(0.01,0)) +
  labs(x = "Normalized lagged dead oyster density",
       y = "Normalized juvenile oyster density") +
  theme_classic()
```

```{r Analyses of juvenile oyster density ~ lagged dead oyster density}

#Check normality of Spat and lagged_Dead
hist(vcr_lagged$Spat)
hist(vcr_lagged$lagged_Dead) # both positively skewed

# Log-transform to increase normality
vcr_lagged$spat.log <- log1p(vcr_lagged$Spat)  # log1p computes log(x + 1)
vcr_lagged$lagged_dead.log <- log1p(vcr_lagged$lagged_Dead)

hist(vcr_lagged$spat.log)
hist(vcr_lagged$lagged_dead.log)

# Simple linear model
model <- lm(spat.log ~ lagged_dead.log, data = vcr_lagged)
summary(model)

# glmm of 
juv_oyst.glmm <- glmmTMB(Spat ~ lagged_Dead + (1|site),
                         data = vcr_lagged)
summary(juv_oyst.glmm)

predictions.juv_oyst <- ggpredict(juv_oyst.glmm, terms = ~lagged_Dead)
plot(predictions.juv_oyst)
predictions.juv_oyst <- as.data.frame(predictions.juv_oyst)

# Model visualization
ggplot()+
  geom_ribbon(data = predictions.juv_oyst, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = predictions.juv_oyst, 
            aes(x = x , y = predicted))+
  geom_point(data = vcr_lagged, 
             aes(x = lagged_Dead, y = Spat),
             alpha = 0.6,
             size = 2) +
  labs(x = Dead~oyster~density~(Ind/0.25~m^2)~prev~yr,
       y = Juvenile~oyster~density~(Ind/0.25~m^2)) +
  theme_classic()

# glmm without summer, due to opposite (negative) relationship in summer
vcr_lagged.fall_spring <- vcr_lagged %>% 
  filter(season != "Summer") %>% 
  group_by(year, site) %>% 
  summarize(agg.juv_density = sum(Spat),
            agg.dead_density = sum(Dead))

juv_oyst.glmm.fall_spring <- glmmTMB(agg.juv_density ~ agg.dead_density + (1|site),
                         data = vcr_lagged.fall_spring)
summary(juv_oyst.glmm.fall_spring)

predictions.juv_oyst.fall_spring <- ggpredict(juv_oyst.glmm.fall_spring, terms = ~agg.dead_density)
plot(predictions.juv_oyst.fall_spring)
predictions.juv_oyst.fall_spring <- as.data.frame(predictions.juv_oyst.fall_spring)

# Model visualization
ggplot()+
  geom_ribbon(data = predictions.juv_oyst.fall_spring, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = predictions.juv_oyst.fall_spring, 
            aes(x = x , y = predicted))+
  geom_point(data = vcr_lagged.fall_spring, 
             aes(x = agg.dead_density, y = agg.juv_density),
             alpha = 0.6,
             size = 2) +
  labs(x = "Dead oyster density",
       y = "Juvenile oyster density") +
  theme_classic()




```

```{r}
## Normalized densities
vcr_lagged.norm <- vcr_lagged.fall_spring %>%
  mutate(dead.norm = agg.dead_density/max(vcr_lagged.fall_spring$agg.dead_density),
         spat.norm = agg.juv_density/max(vcr_lagged.fall_spring$agg.juv_density))

# glmm of normalized juvenile oyster density ~ normalized dead oyster density
juv_oyst.norm.glmm <- glmmTMB(spat.norm ~ dead.norm + (1|site),
                         data = vcr_lagged.norm)
summary(juv_oyst.norm.glmm)

predictions.juv_oyst.norm <- ggpredict(juv_oyst.norm.glmm, terms = ~dead.norm)
plot(predictions.juv_oyst.norm)
predictions.juv_oyst.norm <- as.data.frame(predictions.juv_oyst.norm)

# Model visualization
ggplot()+
  geom_ribbon(data = predictions.juv_oyst.norm, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = predictions.juv_oyst.norm, 
            aes(x = x , y = predicted))+
  geom_point(data = vcr_lagged.norm, 
             aes(x = dead.norm, y = spat.norm),
             alpha = 0.6,
             size = 2) +
  labs(x = "Normalized dead oyster density",
       y = "Normalized juvenile oyster density") +
  scale_x_continuous(expand = c(0,0.02),
                     limits = c(0,1.01)) +
  scale_y_continuous(expand = c(0,0.02),
                     limits = c(0,1.01)) +
  theme_classic()
```


```{r Time-averaged analysis}
# summarize by year
vcr_lagged.summary <- vcr_lagged %>% 
  group_by(year) %>% 
  summarize(mean_dead = mean(lagged_Dead),
            se_dead = sd(lagged_Dead)/sqrt(n()),
            mean_spat = mean(Spat),
            se_spat = sd(Spat)/sqrt(n())) %>% 
  filter(year != "2006")

ggplot(vcr_lagged.summary, aes(x = mean_dead, y = mean_spat)) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmax = mean_dead + se_dead,
                    xmin = mean_dead - se_dead)) +
  geom_errorbar(aes(ymax = mean_spat + se_spat,
                    ymin = mean_spat - se_spat)) +
  labs(x = "Dead oyster density (previous year)",
       y = "Juvenile oyster density") +
  theme_classic()

```

## Harvard Forest
```{r}

hvf <- read_csv("Data/Harvard Forest/hf126-02-tree.csv") %>% 
  clean_names()

```

## H.J. Andrews Experimental Forest
```{r}

# Load in data and filter for observations wiht Study ID 'OHJA'
hja.downed_wood <- read_csv("Data/Andrews Forest/Coarse woody debris/TD01204.csv") %>% 
  clean_names() %>% 
  filter(studyid == "OHJA")

# Summarize deadwood metrics by plot
hja.downed_wood.summary <- hja.downed_wood %>% 
  group_by(stand, plot, year) %>% 
  summarize(total_volume = sum(wvol),
            total_mass = sum(wmass),
            total_cover = sum(pcover),
            total_area = sum(surfarea))

write_csv(hja.downed_wood.summary, "Data/Andrews Forest/Coarse woody debris/OHJA_downed wood summary.csv")

```

## Moorea Coral Reef
```{r}

## Live coral and macroalgae
# Load wide data table
mcr_coral.wide <- read_csv("Data/Moorea Coral Reef/knb-lter-mcr.4_2_20240105.csv") %>% 
  clean_names() %>% 
  dplyr::select(-c(3:6, 21, 34)) 

# Convert to long format
mcr_coral.long <- mcr_coral.wide %>% 
  pivot_longer(cols = c(3:28), names_to = "taxa", values_to = "percent_cover") %>%
  mutate(funct_group = case_when(taxa == "macroalgae" ~ "Macroalgae",
                                 TRUE ~ "Hard coral"),
         percent_cover = replace(percent_cover, percent_cover == "na", 0),
         percent_cover = as.numeric(percent_cover))

mcr <- mcr_coral.long %>% 
  mutate(site = str_sub(location, 0, 6)) %>% 
  filter(str_detect(location, "Outer"))

## Dead coral
mcr.dead <- read_csv("Data/Moorea Coral Reef/MCR_LTER_CoralStruct_Coral_Colony_Desc_2006-2011_20120530.csv") %>% 
  clean_names()

mcr.dead <- mcr.dead %>% 
  filter(alive_or_dead == "Dead") %>% 
  mutate(perc_cover = (area_cm2/2500)*100)

ggplot(mcr.dead, aes(x = year, y = perc_cover, color = site)) +
  geom_point()

ggplot(mcr, aes(x = date, y = perc_cover, color = site)) +
  geom_point()

```

#### Change in live coral cover ~ dead coral cover
```{r Data wrangling for dead coral cover and algae}

# Live and dead coral annotations
regions_master <- read_csv("Data/Moorea Coral Reef/Regions_master.csv") %>% 
  clean_names

# Create dataframe of dead coral cover
dead_coral <- regions_master %>%
  filter(tag_lab_class_name == "Dead coral") %>% 
  group_by(plot, treatment, tag_lab_date) %>% 
  summarize(dead_coral.m_sq = sum(tag_lab_surf_area)*0.0001) %>% 
  mutate(time_point = case_when(tag_lab_date == "8/1/19" ~ 2019,
                                       tag_lab_date == "8/1/20" ~ 2020,
                                       tag_lab_date == "8/1/21" ~ 2021,
                                       tag_lab_date == "8/1/22" ~ 2022,
                                       tag_lab_date == "8/1/23" ~ 2023),
         time_point = as.factor(time_point))

```

#### Annual loss of live coral ~ dead coral cover
```{r Data wrangling for live coral loss ~ algae cover}

# Create dataframe for live coral only
live_coral <- regions_master %>% 
  filter(tag_lab_class_name == "Pocillopora" | tag_lab_class_name == "Acropora") %>% 
  group_by(plot, treatment, tag_lab_date) %>% 
  summarize(live_coral.total_surf = sum(tag_lab_surf_area)) %>% 
  mutate(m_sq = live_coral.total_surf*0.0001,
         years_since_start = case_when(tag_lab_date == "8/1/19" ~ 0,
                                       tag_lab_date == "8/1/20" ~ 1,
                                       tag_lab_date == "8/1/21" ~ 2,
                                       tag_lab_date == "8/1/22" ~ 3,
                                       tag_lab_date == "8/1/23" ~ 4),
         time_point = case_when(tag_lab_date == "8/1/19" ~ 2019,
                                       tag_lab_date == "8/1/20" ~ 2020,
                                       tag_lab_date == "8/1/21" ~ 2021,
                                       tag_lab_date == "8/1/22" ~ 2022,
                                       tag_lab_date == "8/1/23" ~ 2023),
         time_point = as.factor(time_point))


# Create dataframe for proportional changes in live coral in each year (time period)
coral_change.prop <- live_coral %>%
  select(-c(tag_lab_date, live_coral.total_surf, years_since_start)) %>%
  rename(live_coral.m_sq = m_sq) %>%
  pivot_wider(names_from = time_point, values_from = live_coral.m_sq) %>%
  rename(year_0 = "2019",
         year_1 = "2020",
         year_2 = "2021",
         year_3 = "2022",
         year_4 = "2023") %>%
  mutate(delta.1 = (year_1 - year_0)/year_0,
         delta.2 = (year_2 - year_1)/year_1,
         delta.3 = (year_3 - year_2)/year_2,
         delta.4 = (year_4 - year_3)/year_3)

coral_change.prop <- coral_change.prop %>%
  select(-c(year_0, year_1, year_2, year_3, year_4)) %>%
  pivot_longer(cols = c("delta.1", "delta.2", "delta.3", "delta.4"),
               names_to = "delta_year",
               values_to = "delta_coral") %>%
  drop_na() %>%
  mutate(period = case_when(delta_year == "delta.1" ~ "2019 - 2020",
                          delta_year == "delta.2" ~ "2020 - 2021",
                          delta_year == "delta.3" ~ "2021 - 2022",
                          delta_year == "delta.4" ~ "2022 - 2023"))

# Create dataframe for average algae cover in each time period
dead_coral.means <- dead_coral %>%
  select(-tag_lab_date) %>% 
  pivot_wider(names_from = time_point, values_from = dead_coral.m_sq) %>%
  rename(year_0 = "2019",
         year_1 = "2020",
         year_2 = "2021",
         year_3 = "2022",
         year_4 = "2023") %>%
  mutate(mean.1 = (year_1 + year_0)/2,
         mean.2 = (year_2 + year_1)/2,
         mean.3 = (year_3 + year_2)/2,
         mean.4 = (year_4 + year_3)/2)

dead_coral.means <- dead_coral.means %>%
  select(-c(year_0, year_1, year_2, year_3, year_4)) %>%
  pivot_longer(cols = c("mean.1", "mean.2", "mean.3", "mean.4"),
               names_to = "period",
               values_to = "dead.mean") %>%
  drop_na() %>%
  mutate(period = case_when(period == "mean.1" ~ "2019 - 2020",
                          period == "mean.2" ~ "2020 - 2021",
                          period == "mean.3" ~ "2021 - 2022",
                          period == "mean.4" ~ "2022 - 2023"))

live_dead <- merge(coral_change.prop, dead_coral.means)
live_dead <- live_dead %>% 
  mutate(delta_coral.trans = delta_coral*(-1))

ggplot(live_dead, aes(x = dead.mean, y = delta_coral)) +
  geom_point()

```

```{r Stats for change in coral ~ dead coral cover}

# Linear mixed effects model of annual proportional change in coral cover ~ mean algae cover in the same period; time period and plot identity as random effects
live_dead.glmm <- glmmTMB(delta_coral ~ dead.mean + (1|period) + (1|plot), 
                    family = gaussian(link = "identity"),
                    data = live_dead)
summary(live_dead.glmm)

# Get confidence interval for model
confint(live_dead.glmm)

# Check residuals
hist(residuals(live_dead.glmm))
car::qqPlot(residuals(live_dead.glmm))

# Plot model predictions and create dataframe 
plot(ggpredict(live_dead.glmm, terms = ~dead.mean))
live_dead.predictions <- as.data.frame(ggpredict(live_dead.glmm, terms = ~dead.mean))

```

```{r}
live_dead <- live_dead %>% 
  mutate(delta_coral.norm = (delta_coral/-0.4589132)*(-1),
         dead.mean.norm = dead.mean/4.5809965)

live_dead.norm.glmm <- glmmTMB(delta_coral.norm ~ dead.mean.norm + (1|period) + (1|plot), 
                    family = gaussian(link = "identity"),
                    data = live_dead)
summary(live_dead.glmm)

plot(ggpredict(live_dead.norm.glmm, terms = ~dead.mean.norm))
live_dead.predictions.norm <- as.data.frame(ggpredict(live_dead.norm.glmm, terms = ~dead.mean.norm))
```

```{r Visualizaton of proportional coral change ~ mean dead cover in each period}

# Model visualization
live_dead %>%
  ggplot()+
  geom_ribbon(data = live_dead.predictions, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = live_dead.predictions, 
            aes(x = x, y = predicted))+
  geom_point(aes(x = dead.mean, y = delta_coral),
             alpha = 0.6,
             size = 2)+
  labs(x = Mean~annual~dead~coral~cover~(m^2),
       y = "Annual prop. change in live coral cover") +
  # scale_shape_manual(values = c(19,17),
  #                    labels = c("Removal", "Retention"),
  #                    name = "Skeletal treatment") +
  theme_classic(base_size = 12)

# Normalized
live_dead %>%
  ggplot()+
  geom_ribbon(data = live_dead.predictions.norm, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = live_dead.predictions.norm, 
            aes(x = x, y = predicted))+
  geom_point(aes(x = dead.mean.norm, y = delta_coral.norm),
             alpha = 0.6,
             size = 2)+
  labs(x = "Normalized annual dead coral cover",
       y = "Normalized annual change in live coral") +
  scale_x_continuous(expand = c(0,0.02),
                     limits = c(0,1.01)) +
  #scale_y_continuous(expand = c(0,0.02),
                     #limits = c(0,-1.01)) +
  theme_classic(base_size = 10)

```

## Santa Barbara Coastal

Questions:
1. What is the transect/quadrat sampling structure?

2. What would dead holdfasts be likely to influence? Abundance of certain species? If so, which species?

Notes from meeting with Kyle:
Parse out the algae and invert species further
- Look at relationships between dead holdfasts and macrocystis, holdfasts and canopy-forming (macrocystis, cystoseira, egregia, and sargassum), dead holdfasts and  understory algal species (everything else); compare this with all algae grouped together
- For inverts, parse out mobile vs sessile, look at relationship between holdfasts and sessile

- Maybe remove quads with 0% cover of dead holdfasts

```{r}

# sbc <- read_csv("Data/Santa Barbara Coastal/Annual_Cover_All_Years_20240823.csv") %>% 
#   clean_names() %>% 
#   filter(group != "INVERT")
# 
# write_csv(sbc, "Data/Santa Barbara Coastal/Annual_Cover_All_Years_algae-only.csv")

sbc <- read_csv("Data/Santa Barbara Coastal/Annual_Cover_All_Years_algae-only.csv") %>% 
  clean_names() 

# Create dataframe with only dead macrocystis holdfasts
sbc_dead <- sbc %>%
  filter(sp_code == "DMH",
         percent_cover > -1) %>%
  select(year, site, transect, quad, side, percent_cover) %>%
  rename(dmh_cover = percent_cover)

# Look at distribution of dead holdfast cover on different transects over time
sbc_dead %>%
  group_by(year, site, transect, quad) %>%
  summarize(dead_cover = sum(dmh_cover)) %>%
  ggplot(aes(x = dead_cover)) +
    geom_histogram() +
    facet_wrap(~year)

```

```{r}
# Create dataframe of inverts and algae only
sbc.groups <- sbc %>%
  filter(percent_cover > -1,
         sp_code != "DMH",
         group != "PLANT") %>%
  group_by(year, site, transect, quad, group) %>%
  summarize(total_cover = sum(percent_cover))

sbc.merged <- merge(sbc.groups, sbc_dead)

sbc.lagged <- sbc.merged %>%
  mutate(dmh_cover.lagged = lag(dmh_cover))

ggplot(sbc.merged, aes(x = dmh_cover, y = total_cover)) +
  geom_point(alpha = 0.7) +
  facet_wrap(~group) +
  theme_classic()

```

### Live macrocystis cover ~ dead holdfast cover
```{r}
# Dataframe with only live macrocystis and dead holdfasts
sbc_macro <- sbc %>%
  filter(percent_cover > -1,
         scientific_name == "Macrocystis pyrifera") %>% 
  select(year, site, transect, quad, side, percent_cover) %>% 
  rename(macro_cover = percent_cover)

# Merge dead holdfast data and live macro data
sbc.macro_dmh <- merge(sbc_macro, sbc_dead)

ggplot(sbc.macro_dmh, aes(x = dmh_cover, y = macro_cover)) +
  geom_jitter()

# Try removing quadrats with 0% cover of dead holdfasts
sbc.macro_dmh.no_zeroes <- sbc.macro_dmh %>% 
  filter(dmh_cover > 0)

ggplot(sbc.macro_dmh.no_zeroes, aes(x = dmh_cover, y = macro_cover)) +
  geom_jitter()

# Calculate means for site x year combos
sbc.macro_dmh.summary <- sbc.macro_dmh.no_zeroes %>% 
  group_by(site, year) %>% 
  summarize(macro_cover.mean = mean(macro_cover),
            dmh_cover.mean = mean(dmh_cover))

ggplot(sbc.macro_dmh.summary, aes(x = dmh_cover.mean, y = macro_cover.mean)) +
  geom_jitter()

```

Statistics
```{r}

# glmm of macro cover ~ dead holdfast cover
sbc.macro_dmh.glmm <- glmmTMB(macro_cover.mean ~ dmh_cover.mean + (1|site) + (1|year),
                              family = gaussian(link = "identity"),
                              data = sbc.macro_dmh.summary)
summary(sbc.macro_dmh.glmm)

# Check residuals
hist(residuals(sbc.macro_dmh.glmm))
car::qqPlot(residuals(sbc.macro_dmh.glmm))

# Plot model predictions and create dataframe 
plot(ggpredict(sbc.macro_dmh.glmm, terms = ~dmh_cover.mean))
sbc.macro_dmh.predictions <- as.data.frame(ggpredict(sbc.macro_dmh.glmm, terms = ~dmh_cover.mean))

```

```{r}
# Normalized cover
sbc.macro_dmh.summary <- sbc.macro_dmh.summary %>% 
  mutate(macro_cover.norm = macro_cover.mean/max(macro_cover.mean),
         dmh_cover.norm = dmh_cover.mean/max(dmh_cover.mean))

# glmm of macro cover ~ dead holdfast cover
sbc.macro_dmh.glmm.norm <- glmmTMB(macro_cover.norm ~ dmh_cover.norm + (1|site) + (1|year),
                              family = gaussian(link = "identity"),
                              data = sbc.macro_dmh.summary)
summary(sbc.macro_dmh.glmm.norm)

# Check residuals
hist(residuals(sbc.macro_dmh.glmm.norm))
car::qqPlot(residuals(sbc.macro_dmh.glmm.norm))

# Plot model predictions and create dataframe 
plot(ggpredict(sbc.macro_dmh.glmm.norm, terms = ~dmh_cover.norm))
sbc.macro_dmh.predictions.norm <- as.data.frame(ggpredict(sbc.macro_dmh.glmm.norm, terms = ~dmh_cover.norm))

```

Visualization of normalized model
```{r}
sbc.macro_dmh.summary %>%
  ggplot()+
  geom_ribbon(data = sbc.macro_dmh.predictions.norm, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = sbc.macro_dmh.predictions.norm, 
            aes(x = x, y = predicted))+
  geom_jitter(aes(x = dmh_cover.norm, y = macro_cover.norm),
             alpha = 0.6,
             size = 2)+
  labs(x = "Normalized % cover of dead holdfasts",
       y = "Normalized % cover of live Macrocystis") +
  theme_classic(base_size = 12)
```

### Canopy-forming species cover ~ dead holdfast cover
```{r}
# Dataframe with only canopy-forming species and dead holdfasts
sbc_canopy <- sbc %>%
  filter(percent_cover > -1,
         taxon_genus %in% c("Macrocystis", "Cystoseira", "Sargassum", "Egregia")) %>% 
  select(year, site, transect, quad, side, percent_cover) %>% 
  rename(canopy_cover = percent_cover)

# Merge dead holdfast data and live macro data and remove quads with 0% dead holdfast cover
sbc.canopy_dmh <- merge(sbc_canopy, sbc_dead)
sbc.canopy_dmh <- sbc.canopy_dmh %>% 
  filter(dmh_cover > 0)

ggplot(sbc.canopy_dmh, aes(x = dmh_cover, y = canopy_cover)) +
  geom_jitter()

# summarize by year & site; calculate normalized values
sbc.canopy_dmh.summary <- sbc.canopy_dmh %>% 
  group_by(site, year) %>% 
  summarize(canopy_cover.mean = mean(canopy_cover),
            dmh_cover.mean = mean(dmh_cover)) %>% 
  mutate(canopy_cover.norm = canopy_cover.mean/max(canopy_cover.mean),
         dmh_cover.norm = dmh_cover.mean/max(dmh_cover.mean))

```

```{r}
# glmm of macro cover ~ dead holdfast cover
sbc.canopy_dmh.glmm.norm <- glmmTMB(canopy_cover.norm ~ dmh_cover.norm + (1|site) + (1|year),
                              family = gaussian(link = "identity"),
                              data = sbc.canopy_dmh.summary)
summary(sbc.canopy_dmh.glmm.norm)

# Check residuals
hist(residuals(sbc.canopy_dmh.glmm.norm))
car::qqPlot(residuals(sbc.canopy_dmh.glmm.norm))

# Plot model predictions and create dataframe 
plot(ggpredict(sbc.canopy_dmh.glmm.norm, terms = ~dmh_cover.norm))
sbc.canopy_dmh.predictions.norm <- as.data.frame(ggpredict(sbc.canopy_dmh.glmm.norm, terms = ~dmh_cover.norm))

```

Visualization of normalized model
```{r}
sbc.canopy_dmh.summary %>%
  ggplot()+
  geom_ribbon(data = sbc.canopy_dmh.predictions.norm, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = sbc.canopy_dmh.predictions.norm, 
            aes(x = x, y = predicted))+
  geom_jitter(aes(x = dmh_cover.norm, y = canopy_cover.norm),
             alpha = 0.6,
             size = 2)+
  labs(x = "Normalized % cover of dead holdfasts",
       y = "Normalized % cover of canopy-forming spp.") +
  theme_classic(base_size = 12)
```

### Understory algal spp cover ~ dead holdfast cover
```{r}
# Dataframe with only understory species and dead holdfasts
sbc_under <- sbc %>%
  filter(percent_cover > -1,
         group == "ALGAE",
         taxon_genus != "Macrocystis", 
         taxon_genus !="Cystoseira", 
         taxon_genus != "Sargassum", 
         taxon_genus !="Egregia") %>% 
  select(year, site, transect, quad, side, percent_cover) %>% 
  rename(under_cover = percent_cover)

# Merge dead holdfast data and live macro data and remove quads with 0% dead holdfast cover
sbc.under_dmh <- merge(sbc_under, sbc_dead)
sbc.under_dmh <- sbc.under_dmh %>% 
  filter(dmh_cover > 0)

ggplot(sbc.under_dmh, aes(x = dmh_cover, y = under_cover)) +
  geom_jitter()

# summarize by year & site; calculate normalized values
sbc.under_dmh.summary <- sbc.under_dmh %>% 
  group_by(site, year) %>% 
  summarize(under_cover.mean = mean(under_cover),
            dmh_cover.mean = mean(dmh_cover)) %>% 
  mutate(under_cover.norm = under_cover.mean/max(under_cover.mean),
         dmh_cover.norm = dmh_cover.mean/max(dmh_cover.mean))

```

```{r}
# glmm of macro cover ~ dead holdfast cover
sbc.under_dmh.glmm.norm <- glmmTMB(under_cover.norm ~ dmh_cover.norm + (1|site) + (1|year),
                              family = gaussian(link = "identity"),
                              data = sbc.under_dmh.summary)
summary(sbc.under_dmh.glmm.norm)

# Check residuals
hist(residuals(sbc.under_dmh.glmm.norm))
car::qqPlot(residuals(sbc.under_dmh.glmm.norm))

# Plot model predictions and create dataframe 
plot(ggpredict(sbc.under_dmh.glmm.norm, terms = ~dmh_cover.norm))
sbc.under_dmh.predictions.norm <- as.data.frame(ggpredict(sbc.under_dmh.glmm.norm, terms = ~dmh_cover.norm))

```

Visualization of normalized model
```{r}
sbc.under_dmh.summary %>%
  ggplot()+
  geom_ribbon(data = sbc.under_dmh.predictions.norm, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = sbc.under_dmh.predictions.norm, 
            aes(x = x, y = predicted))+
  geom_jitter(aes(x = dmh_cover.norm, y = under_cover.norm),
             alpha = 0.6,
             size = 2)+
  labs(x = "Normalized % cover of dead holdfasts",
       y = "Normalized % cover of understory algal spp.") +
  theme_classic(base_size = 12)
```

## Georgia Coastal Ecosystems
```{r}

gce.wrack_disturbance <- read_csv("Data/Georgia Coastal Ecosystems/PLT-GCEM-1801_9_0.CSV") %>% 
  clean_names()

```


## Bonanza Creek
```{r}

bnz_seeds <- read_csv("Data/Bonanza Creek/AK2004 sites seeds.csv") %>% 
  clean_names() %>% 
  select(burn, site, bs_stg_ba, total_m2) %>% 
  mutate(log_ba = log(bs_stg_ba + 0.1), # add small constants to avoid taking log of 0
         log_total_seeds = log(total_m2 + 0.1)) %>% 
    drop_na()

hist(bnz_seeds$log_ba)
hist(bnz_seeds$log_total_seeds)

ggplot(bnz_seeds, aes(x = log_ba, y = log_total_seeds)) +
  geom_point() +
  #geom_smooth(method = "lm") +
  theme_classic()

```

Statistics for BNZ seed data
```{r}

bnz_glmm <- glmmTMB(log_total_seeds ~ log_ba + (1|burn),
                    family = gaussian(link = "identity"),
                    data = bnz_seeds)

summary(bnz_glmm)

# Check residuals
hist(residuals(bnz_glmm))
car::qqPlot(residuals(bnz_glmm))

# Plot model predictions and create dataframe 
plot(ggpredict(bnz_glmm, terms = ~log_ba))
bnz_seeds.predictions <- as.data.frame(ggpredict(bnz_glmm, terms = ~log_ba))

```

Model visualization
```{r}
bnz_seeds %>%
  ggplot()+
  geom_ribbon(data = bnz_seeds.predictions, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = bnz_seeds.predictions, 
            aes(x = x, y = predicted))+
  geom_point(aes(x = log_ba, y = log_total_seeds),
             alpha = 0.6,
             size = 2)+
  labs(x = "Log standing basal area of Spruce (m^2/ha)",
       y = "Log seed density (#/m^2)") +
  theme_classic(base_size = 12)
```

Normalized model
```{r}

# Scale predictor and response variable to their respective maximum values
bnz_seeds <- bnz_seeds %>% 
  mutate(log_ba.norm = (log_ba/max(log_ba)),
         log_seeds.norm = log_total_seeds/max(log_total_seeds))

bnz_seeds.norm.glmm <- glmmTMB(log_seeds.norm ~ log_ba.norm + (1|burn),
                    family = gaussian(link = "identity"),
                    data = bnz_seeds)
summary(bnz_seeds.norm.glmm)

plot(ggpredict(bnz_seeds.norm.glmm, terms = ~log_ba.norm))
bnz_seeds.predictions.norm <- as.data.frame(ggpredict(bnz_seeds.norm.glmm, terms = ~log_ba.norm))
```

Model visualization
```{r}
bnz_seeds %>%
  ggplot()+
  geom_ribbon(data = bnz_seeds.predictions.norm, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = bnz_seeds.predictions.norm, 
            aes(x = x, y = predicted))+
  geom_point(aes(x = log_ba.norm, y = log_seeds.norm),
             alpha = 0.6,
             size = 2)+
  labs(x = "Normalized standing basal area of Spruce",
       y = "Normalized seed density") +
  theme_classic(base_size = 12)
```

## Luquillo

```{r}
# Load and clean seedling data, calculate seedling volumes
luq.seedlings <- read_csv("Data/Luquillo/CTESeedlingMeasurementData2003-20212.csv") %>% 
  clean_names() %>% 
  #select(-c(10:14)) %>% 
  #drop_na() %>% 
  mutate(start_date = as.character(start_date),
         year = as.numeric(substr(start_date, 1, 4)))
         #seedling_ba = (pi/4)*diameter,
         #seedling_count = 1) 

luq.litter <- read_csv("Data/Luquillo/CTE-Litterfall-data.csv") %>% 
  clean_names() %>% 
  select(-c(8:10)) %>% 
  mutate(date = as.character(date),
         year = as.numeric(substr(date, 1, 4)),
         litter_mass = leaves_in_g + wood_in_g + misc_in_g + fruits_seeds_flower_g) 

luq.cte_ttt <- read_csv("Data/Luquillo/CTE_Treatments_0.csv") %>% 
  clean_names() %>% 
  select(-c(latitude, longitude, subplot)) %>% 
  distinct()
  
```

```{r}
# Seedling mortality
luq.seedlings_mortality <- luq.seedlings %>% 
  filter(#year %in% c(2004, 2005, 2006, 2007, 2008),
         #year == 2004 | year == 2005,
         dead_alive_notfound != "NF") %>% 
  mutate(mortality = if_else(dead_alive_notfound == "D", 1, 0),
         survival = if_else(dead_alive_notfound == "A", 1, 0),
         year = as.numeric(year))

luq.seedlings_mortality <- merge(luq.seedlings_mortality, luq.cte_ttt)

luq.seedlings_mortality <- luq.seedlings_mortality %>% 
  relocate(treatment, .before = block) %>% 
  #filter(height != "NA") %>% 
  select(-c(lessthan10cm, comments, new))

luq.mortality_counts <- luq.seedlings_mortality %>%   
  group_by(treatment, year, block, plot) %>% 
  summarize(mortality.count = sum(mortality),
            seedling.count = sum(survival))

# Summary stats  
mortality.summary <- luq.mortality_counts %>% 
  group_by(treatment, year) %>% 
  summarize(mortality.mean = mean(mortality.count),
            mortality.se = sd(mortality.count)/sqrt(n()),
            survival.mean = mean(seedling.count),
            survival.se = sd(seedling.count)/sqrt(n()))

ggplot(mortality.summary, aes(x = year, mortality.mean, color = treatment)) +
  geom_line(aes(group = treatment,
                color = treatment)) +
  geom_errorbar(aes(ymax = mortality.mean + mortality.se,
                    ymin = mortality.mean - mortality.se),
                width = 0) +
  geom_point(aes(color = treatment)) +
  labs(x = "Year",
       y = "Seedling mortality (#/yr)") +
  theme_classic()

ggplot(mortality.summary, aes(x = year, survival.mean, color = treatment)) +
  geom_line(aes(group = treatment,
                color = treatment)) +
  geom_errorbar(aes(ymax = survival.mean + survival.se,
                    ymin = survival.mean - survival.se),
                width = 0) +
  geom_point(aes(color = treatment)) +
  labs(x = "Year",
       y = "Seedling count") +
  theme_classic()
  
```

Mixed effects model, seedling count ~ ttt
```{r}
# Remove first year of experiment (pre-manipulation)
luq.seeding_counts <- luq.mortality_counts %>% 
  filter(year %in% c(2005, 2006, 2007, 2008),
         treatment == "Control" | treatment == "Trim+Debris") %>% 
  mutate(year = as.factor(year))
  
# Model 
luq.seedling_counts.glmm <- glmmTMB(seedling.count ~ treatment + year + (1|block),
                                          family = "gaussian", 
                                          data = luq.seeding_counts)

# Check residuals
hist(residuals(luq.seedling_counts.glmm))
car::qqPlot(residuals(luq.seedling_counts.glmm))

summary(luq.seedling_counts.glmm)

# Plot model predictions and create dataframe 
plot(ggpredict(luq.seedling_counts.glmm, terms = ~treatment))
luq.seedlings.predictions <- as.data.frame(ggpredict(luq.seedling_counts.glmm, terms = ~treatment))

```

Model visualization
```{r}
ggplot(luq.seedlings.predictions, aes(x = x, y = predicted))+
  geom_errorbar(aes(ymax = conf.high,
                ymin = conf.low),
                width = 0) +
  geom_point(size = 3) +
  labs(x = "Treatment",
       y = "Seedling count") +
  theme_classic(base_size = 12)


```

Just trimming treatments
```{r}
# Filter for only trim treatments, create normalized variable
luq.seeding_counts.trim <- luq.mortality_counts %>% 
  filter(#treatment == "Control" | treatment == "Trim+Debris",
         year %in% 2009) %>% 
  mutate(treatment = as.factor(treatment),
         #year = as.numeric(year),
         seedling_count.norm = seedling.count/max(luq.mortality_counts$seedling.count))
  
# Model 
luq.seedling_counts.trim.glmm <- glmmTMB(seedling.count ~ treatment + (1|year) + (1|block),
                                          family = "gaussian", 
                                          data = luq.seeding_counts.trim)

# Check residuals
hist(residuals(luq.seedling_counts.trim.glmm))
car::qqPlot(residuals(luq.seedling_counts.trim.glmm))

summary(luq.seedling_counts.trim.glmm)

# Plot model predictions and create dataframe 
plot(ggpredict(luq.seedling_counts.trim.glmm, terms = ~treatment))
luq.seedlings.trim.predictions <- as.data.frame(ggpredict(luq.seedling_counts.trim.glmm, terms = ~treatment))


## Normalized model 
luq.seedling_counts.trim.norm.glmm <- glmmTMB(seedling_count.norm ~ treatment + (1|year) + (1|block),
                                          family = "gaussian", 
                                          data = luq.seeding_counts.trim)

# Check residuals
hist(residuals(luq.seedling_counts.trim.norm.glmm))
car::qqPlot(residuals(luq.seedling_counts.trim.norm.glmm))

summary(luq.seedling_counts.trim.norm.glmm)

# Plot model predictions and create dataframe 
plot(ggpredict(luq.seedling_counts.trim.norm.glmm, terms = ~treatment))
luq.seedlings.trim.predictions.norm <- as.data.frame(ggpredict(luq.seedling_counts.trim.norm.glmm, terms = ~treatment))

```

Model visualizations
```{r}
# Raw data
ggplot(luq.seedlings.trim.predictions, aes(x = x, y = predicted)) +
  geom_point(data = luq.seeding_counts.trim,
             aes(x = treatment, y = seedling.count),
             alpha = 0.6) +
  geom_errorbar(aes(ymax = conf.high,
                ymin = conf.low),
                width = 0) +
  geom_point(size = 3) +
  #scale_x_discrete(labels = c("Litter removed", "Litter added")) +
  scale_y_continuous(limits = c(0,700),
                     expand = c(0,0)) +
  labs(x = "Treatment",
       y = "Seedling count") +
  theme_classic(base_size = 12)

# Normalized data
ggplot(luq.seedlings.trim.predictions.norm, aes(x = x, y = predicted)) +
  geom_point(data = luq.seeding_counts.trim,
             aes(x = treatment, y = seedling_count.norm),
             alpha = 0.6) +
  geom_errorbar(aes(ymax = conf.high,
                ymin = conf.low),
                width = 0) +
  geom_point(size = 3) +
  scale_x_discrete(labels = c("Litter removed", "Litter added")) +
  scale_y_continuous(limits = c(0,1.1),
                     expand = c(0,0)) +
  labs(x = "Treatment",
       y = "Seedling count (normalized)") +
  theme_classic(base_size = 12)

```





First canopy trimming experiment only
```{r}
# Summarize seedling volumes by year>block>plot for first canopy trimming experiment (2003-2013)
luq.seedlings.summary <- luq.seedlings %>% 
  filter(year < 2014) %>% 
  group_by(year, block, plot) %>% 
  summarize(seedling_ba.total = sum(seedling_ba),
            seedling_count.total = sum(seedling_count),
            seedling_dens = seedling_count.total/111) # 111 is the total area (m^2) in which seedling counts were done in each plot (?)

# Summarize litter masses by year>block>plot for first canopy trimming experiment (2003-2013)
luq.litter.summary <- luq.litter %>% 
  filter(year < 2014,
         year != 2002) %>% 
  group_by(year, block, plot) %>% 
  summarize(litter_mass.total = sum(litter_mass))

# Merge seedling, litter, and plot treatment datasets
luq.seedlings_litter <- merge(luq.seedlings.summary, luq.litter.summary)
luq.seedlings_litter <- merge(luq.seedlings_litter, luq.cte_ttt)

# Remove large outlier
luq.seedlings_litter <- luq.seedlings_litter %>% 
  mutate(plot = as.factor(plot),
         year = as.factor(year)) %>% 
  #filter(seedling_ba.total < 400) %>% 
  drop_na()

# Exploratory visualizations
ggplot(luq.seedlings_litter, aes(x = litter_mass.total, y = seedling_count.total, color = treatment)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "Annual seedling basal area (cm^2)",
       x = "Annual litter mass (g)") +
  #scale_y_continuous(limits = c(-1000,200000)) +
  theme_classic()

ttt_summary <- luq.seedlings_litter %>% 
  group_by(treatment, year) %>% 
  summarize(seedling_count.mean = mean(seedling_count.total),
            seedling_count.se = sd(seedling_count.total)/sqrt(n()),
            litter_mass.mean = mean(litter_mass.total),
            litter_mass.se = sd(litter_mass.total)/sqrt(n()))

ggplot(ttt_summary, aes(x = year, y = seedling_count.mean)) +
  geom_line(aes(group = treatment,
                color = treatment)) +
  geom_errorbar(aes(ymax = seedling_count.mean + seedling_count.se,
                    ymin = seedling_count.mean - seedling_count.se,
                    color = treatment),
                width = 0) +
  geom_point(aes(color = treatment)) +
  labs(x = "Year",
       y = "Seedling counts (mean +/- SE") +
  theme_classic()

ggplot(ttt_summary, aes(x = year, y = litter_mass.mean)) +
  geom_line(aes(group = treatment,
                color = treatment)) +
  geom_errorbar(aes(ymax = litter_mass.mean + litter_mass.se,
                    ymin = litter_mass.mean - litter_mass.se,
                    color = treatment),
                width = 0) +
  geom_point(aes(color = treatment)) +
  labs(x = "Year",
       y = "Litter mass (mean +/- SE") +
  theme_classic()

```

```{r}
## Mixed effects model with both canopy clearing treatments
luq.trim <- luq.seedlings_litter %>% 
  filter(treatment == "Trim+Debris" | treatment == "Trim&clear")

luq.trim.glmm <- glmmTMB(log(seedling_count.total) ~ log(litter_mass.total)*treatment + (1|block),
                                          family = "gaussian",
                                          data = luq.trim)

summary(luq.trim.glmm)

# Check residuals
hist(residuals(luq.trim.glmm))
car::qqPlot(residuals(luq.trim.glmm))

# Plot model predictions and create dataframe 
plot(ggpredict(luq.trim.glmm, terms = ~log(litter_mass.total)*treatment))
luq.seedlings_litter.predictions <- as.data.frame(ggpredict(luq.trim.glmm, terms = ~log(litter_mass.total)*treatment))
```

Model visualization
```{r}
luq.seedlings_litter.norm %>%
  ggplot()+
  geom_ribbon(data = luq.seedlings_litter.predictions, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = luq.seedlings_litter.predictions, 
            aes(x = x, y = predicted))+
  geom_point(aes(x = litter.norm, y = seedlings.norm),
             alpha = 0.6,
             size = 2)+
  labs(x = "Normalized litter biomass",
       y = "Normalized seedling volume") +
  theme_classic(base_size = 12)


```

Mixed effects models of seedling basal area ~ litter mass
```{r}
## Raw data
luq.cte.glmm.all_ttt <- glmmTMB(seedling_ba.total ~ litter_mass.total + treatment + (1|block),
                                          family = "gaussian", 
                                          data = luq.seedlings_litter)

# Check residuals
hist(residuals(luq.cte.glmm.all_ttt))
car::qqPlot(residuals(luq.cte.glmm.all_ttt))

summary(luq.cte.glmm.all_ttt)

# Plot model predictions and create dataframe 
plot(ggpredict(luq.cte.glmm.all_ttt, terms = ~litter_mass.total + treatment))
luq.cte.predictions.all_ttt <- as.data.frame(ggpredict(luq.cte.glmm.all_ttt, terms = ~litter_mass.total + treatment))

## Log-transformed data, basal area

luq.all_ttt.glmm <- glmmTMB(log(seedling_count.total) ~ log(litter_mass.total) + treatment + (1|block),
                                          family = "gaussian", 
                                          data = luq.seedlings_litter)

summary(luq.all_ttt.glmm)

# Check residuals
hist(residuals(luq.all_ttt.glmm))
car::qqPlot(residuals(luq.all_ttt.glmm))

# Plot model predictions and create dataframe 
plot(ggpredict(luq.all_ttt.glmm, terms = ~log(litter_mass.total) + treatment))
luq.cte.predictions.all_ttt.log <- as.data.frame(ggpredict(luq.all_ttt.glmm, terms = ~litter_mass.total))

## Log-transformed data, seedling density
luq.cte.glmm.all_ttt.log.dens <- glmmTMB(log(seedling_dens) ~ log(litter_mass.total) + treatment + (1|block),
                                          family = "gaussian", 
                                          data = luq.seedlings_litter)

summary(luq.cte.glmm.all_ttt.log.dens)

# Check residuals
hist(residuals(luq.cte.glmm.all_ttt.log.dens))
car::qqPlot(residuals(luq.cte.glmm.all_ttt.log.dens))

# Plot model predictions and create dataframe 
plot(ggpredict(luq.cte.glmm.all_ttt.log.dens, terms = ~litter_mass.total))
luq.cte.predictions.all_ttt.log.dens <- as.data.frame(ggpredict(luq.cte.glmm.all_ttt.log.dens, terms = ~litter_mass.total))
```

Model visualization
```{r}
luq.seedlings_litter %>%
  ggplot()+
  geom_ribbon(data = luq.cte.predictions.all_ttt, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high, group = group, fill = group), 
              alpha =0.5)+
  geom_line(data = luq.cte.predictions.all_ttt, 
            aes(x = x, y = predicted, group = group, color = group))+
  geom_point(aes(x = litter_mass.total, y = seedling_vol.total, color = treatment),
             alpha = 0.6,
             size = 2) +
  labs(x = "Litter biomass (g)",
       y = "Seedling volume (cm^2)") +
  scale_x_continuous(limits = c(0,3000)) +
  theme_classic(base_size = 12)

# Log-transformed, basal area
luq.seedlings_litter %>%
  ggplot()+
  geom_ribbon(data = luq.cte.predictions.all_ttt.log, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5)+
  geom_line(data = luq.cte.predictions.all_ttt.log, 
            aes(x = x, y = predicted, group = group))+
  geom_point(aes(x = litter_mass.total, y = seedling_ba.total),
             alpha = 0.6,
             size = 2) +
  labs(x = "Log litter biomass (g)",
       y = "Log seedling basal area (cm^2)") +
  #scale_x_continuous(limits = c(0,3000)) +
  theme_classic(base_size = 12)

# Seedling density
luq.seedlings_litter %>%
  ggplot()+
  geom_ribbon(data = luq.cte.predictions.all_ttt.log.dens, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5)+
  geom_line(data = luq.cte.predictions.all_ttt.log.dens, 
            aes(x = x, y = predicted))+
  geom_point(aes(x = litter_mass.total, y = seedling_dens),
             alpha = 0.6,
             size = 2) +
  labs(x = "Log litter biomass (g)",
       y = "Log seedling density (ind/m^2)") +
  #scale_x_continuous(limits = c(0,3000)) +
  theme_classic(base_size = 12)

```

Normalized data
```{r}
# Scale variables to maximum values
luq.seedlings_litter.norm <- luq.seedlings_litter %>%
  drop_na() %>% 
  mutate(seedlings.norm = seedling_vol.total/max(seedling_vol.total),
         litter.norm = litter_mass.total/max(litter_mass.total))

# Mixed effects model of seedling volume ~ litter mass, normalized
luq.cte.glmm.norm.all_ttt <- glmmTMB(seedlings.norm ~ litter.norm*treatment + (1|block),
                                          family = "gaussian", 
                                          data = luq.seedlings_litter.norm)

summary(luq.cte.glmm.norm.all_ttt)

# Check residuals
hist(residuals(luq.cte.glmm.norm.all_ttt))
car::qqPlot(residuals(luq.cte.glmm.norm.all_ttt))

# Plot model predictions and create dataframe 
plot(ggpredict(luq.cte.glmm.norm, terms = ~litter.norm))
luq.cte.predictions.norm <- as.data.frame(ggpredict(luq.cte.glmm.norm, terms = ~litter.norm))
```

Normalized model visualization
```{r}
luq.seedlings_litter.norm %>%
  ggplot()+
  geom_ribbon(data = luq.cte.predictions.norm, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = luq.cte.predictions.norm, 
            aes(x = x, y = predicted))+
  geom_point(aes(x = litter.norm, y = seedlings.norm),
             alpha = 0.6,
             size = 2)+
  labs(x = "Normalized litter biomass",
       y = "Normalized seedling volume") +
  theme_classic(base_size = 12)
```

Mixed effects models of seedling volume ~ litter mass, without 'No trim + debris' TTT
```{r}
## Raw data
# Remove "Trim + Debris" treatment and scale variables to max values
luq.seedlings_litter.norm <- luq.seedlings_litter %>% 
  filter(treatment != "NoTrim+Debris") %>% 
  drop_na() %>% 
  mutate(seedlings.norm = seedling_vol.total/max(seedling_vol.total),
         litter.norm = litter_mass.total/max(litter_mass.total))

## Mixed effects model of seedling volume ~ litter mass, normalized
luq.cte.glmm <- glmmTMB(seedling_vol.total ~ litter_mass.total + (1|treatment) + (1|block),
                                          family = "gaussian", 
                                          data = luq.seedlings_litter.norm)

# Check residuals
hist(residuals(luq.cte.glmm))
car::qqPlot(residuals(luq.cte.glmm))

# Plot model predictions and create dataframe 
plot(ggpredict(luq.cte.glmm, terms = ~litter_mass.total*treatment))
luq.cte.predictions <- as.data.frame(ggpredict(luq.cte.glmm, terms = ~litter_mass.total*treatment))
```

Model visualization
```{r}
luq.seedlings_litter.norm %>%
  ggplot()+
  geom_ribbon(data = luq.cte.predictions, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high, group = group, fill = group), 
              alpha =0.5)+
  geom_line(data = luq.cte.predictions, 
            aes(x = x, y = predicted, group = group, color = group))+
  geom_point(aes(x = litter_mass.total, y = seedling_vol.total, color = treatment),
             alpha = 0.6,
             size = 2)+
  labs(x = "Litter biomass",
       y = "Seedling volume") +
  theme_classic(base_size = 12)

```


```{r}
## Normalized data
# Mixed effects model of seedling volume ~ litter mass, normalized
luq.cte.glmm.norm <- glmmTMB(seedlings.norm ~ litter.norm + (1|block),
                                          family = "gaussian", # Need to figure out which family to use
                                          data = luq.seedlings_litter.norm)

# Check residuals
hist(residuals(luq.cte.glmm.norm))
car::qqPlot(residuals(luq.cte.glmm.norm))

# Plot model predictions and create dataframe 
plot(ggpredict(luq.cte.glmm.norm, terms = ~litter.norm))
luq.cte.predictions.norm <- as.data.frame(ggpredict(luq.cte.glmm.norm, terms = ~litter.norm))
```

Normalized model visualization
```{r}
luq.seedlings_litter.norm %>%
  ggplot()+
  geom_ribbon(data = luq.cte.predictions.norm, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = luq.cte.predictions.norm, 
            aes(x = x, y = predicted))+
  geom_point(aes(x = litter.norm, y = seedlings.norm),
             alpha = 0.6,
             size = 2)+
  labs(x = "Normalized litter biomass",
       y = "Normalized seedling volume") +
  theme_classic(base_size = 12)
```



Mixed effects models of seedling volume ~ litter mass, by treatment
```{r}
# Remove "Trim + Debris" treatment and scale variables to max values
luq.seedlings_litter.norm <- luq.seedlings_litter %>% 
  filter(treatment != "Trim+Debris") %>% 
  drop_na() %>% 
  mutate(seedlings.norm = seedling_vol.total/max(seedling_vol.total),
         litter.norm = litter_mass.total/max(litter_mass.total))

## Mixed effects model of 'Control' ttt
luq.cte_control <- luq.seedlings_litter.norm %>% 
  filter(treatment == "Control")

luq.cte_control.glmm <- glmmTMB(seedlings.norm ~ litter.norm + (1|block),
                                          family = beta_family(link = "logit"), # Need to figure out which family to use
                                          data = luq.cte_control)

# Check residuals
hist(residuals(luq.cte_control.glmm))
car::qqPlot(residuals(luq.cte_control.glmm))


## Mixed effects model of 'Trim + Clear' ttt
luq.trim_clear <- luq.seedlings_litter.norm %>% 
  filter(treatment == "Trim&clear")

luq.trim_clear.glmm <- glmmTMB(seedlings.norm ~ litter.norm + (1|block),
                                          family = gaussian(link = "identity"),
                                          data = luq.trim_clear)

# Check residuals
hist(residuals(luq.trim_clear.glmm))
car::qqPlot(residuals(luq.trim_clear.glmm))

## Mixed effects model of 'Trim + Clear' ttt
luq.NoTrim_debris <- luq.seedlings_litter.norm %>% 
  filter(treatment == "NoTrim+Debris")

luq.NoTrim_debris.glmm <- glmmTMB(seedlings.norm ~ litter.norm + (1|block),
                                          family = gaussian(link = "identity"),
                                          data = luq.NoTrim_debris)

# Check residuals
hist(residuals(luq.NoTrim_debris.glmm))
car::qqPlot(residuals(luq.NoTrim_debris.glmm))


```

```{r}
# Plot model predictions and create dataframe 
plot(ggpredict(luq.seedlings_litter.norm.glmm, terms = ~litter.norm))
luq.seedlings_litter.predictions <- as.data.frame(ggpredict(luq.seedlings_litter.norm.glmm, terms = ~litter.norm))
```

Model visualization
```{r}
luq.seedlings_litter.norm %>%
  ggplot()+
  geom_ribbon(data = luq.seedlings_litter.predictions, 
              aes(x = x, y = predicted, ymin =conf.low, ymax = conf.high), 
              alpha =0.5,
              fill = "#3B4F8E")+
  geom_line(data = luq.seedlings_litter.predictions, 
            aes(x = x, y = predicted))+
  geom_point(aes(x = litter.norm, y = seedlings.norm),
             alpha = 0.6,
             size = 2)+
  labs(x = "Normalized litter biomass",
       y = "Normalized seedling volume") +
  theme_classic(base_size = 12)
```

# Analyses across systems
## Effect sizes
```{r}

library(glmmTMB)
library(broom.mixed)

# Extract effect sizes from each mixed model
knz.effects <- broom.mixed::tidy(knz.norm.glmm, effects = "fixed", conf.int = TRUE)
  knz.effects[knz.effects$term == "predictor", ]
  
vcr.effects <- broom.mixed::tidy(juv_oyst.norm.glmm, effects = "fixed", conf.int = TRUE)
  vcr.effects[vcr.effects$term == "predictor", ]

mcr.effects <- broom.mixed::tidy(live_dead.norm.glmm, effects = "fixed", conf.int = TRUE)
  mcr.effects[mcr.effects$term == "predictor", ]
  
bnz.effects <- broom.mixed::tidy(bnz_seeds.norm.glmm, effects = "fixed", conf.int = TRUE)
  bnz.effects[bnz.effects$term == "predictor", ]
  
sbc.effects <- broom.mixed::tidy(sbc.canopy_dmh.glmm.norm, effects = "fixed", conf.int = TRUE)
  
luq.effects <- broom.mixed::tidy(luq.seedling_counts.trim.norm.glmm, effects = "fixed", conf.int = TRUE)


# Assemble effect sizes from each ecosystem type
lter.effects <- rbind(knz.effects, vcr.effects, mcr.effects, bnz.effects, sbc.effects, luq.effects)
lter.effects <- lter.effects %>% 
  filter(term != "(Intercept)") %>% 
  mutate(ecosystem = c("Tallgrass prairie", "Oyster reef", "Coral reef", "Boreal forest", "Kelp forest", "Tropical forest"),
         significance = ifelse(conf.low > 0 | conf.high < 0, "Significant", "Not Significant"))
```

```{r}
# Effect size plot
ggplot(lter.effects, aes(x = estimate, y = reorder(ecosystem, -estimate))) + #, color = significance)) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 height = 0,
                 linewidth = 1) +  
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Effect of material legacy on demographic metric ( 95% CI)",
    y = "Ecosystem type") +
  # scale_color_manual(name = "",
  #                    labels = c("Significant", "Non-significant"),
  #                    values = c("black", "grey")) +
  theme_minimal()
  
```

